Formalizing Stack Safety as a Security Property

### 2022-10-05 submission5/noexceptions.tex
Bunch of stuff

\section{Threat Model, Machines, and Enforcement }

Before introducing our formal model of stack safety, we describe
our machine exemplar, enforcement mechanism, and threat model.

We begin with a RISC-V-like machine enhanced with PIPE, a tag-based reference
monitor. We extend this machine with a {\it security semantics},
which divides a program into dynamic {\it principals} that ought to be
protected from one another, and specifies that protection over a set
of abstract {\it security operations}. Specifications may take the form of
predicates on future states: statements of the form,
``When control returns to me...'', or of relations on traces of future execution
(hyper-properties.)

As opposed to a fully-abstract overlay semantics, in which overlay operations
are given safe-by-construction semantics then proven fully abstract,
our machine is not safe unless the monitor is correctly instantiated;
the security semantics gives us the means to distinguish a monitor that implements
stack safety from one that does not.

The underlying machine is a standard RISC-V ISA as modeled in [MIT-RISCV].
Our overlay semantics provide overlay operations for the following code features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers according
  to the RISC-V ABI
\item Arguments passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

@@

We trust the compiler has correctly placed macro-instructions according to
the semantics of the source language, but we do not assume that the code
is otherwise reasonable. In particular, while we are agnostic as to the source
language, C is very plausible, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code.

In general, it is impossible to distinguish buggy source code from an attacker;
we assume nothing about the intent of the source code, except that it makes calls
and returns, and we must guarantee that the caller and callee are protected from one
another.

@@

We do not support dynamic code generation, and our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
We model memory safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

\subsection{Security Semantics}

We begin with a machine that supports a number of security operations,
defined above. For purposes of specifying security, we do not need to
know the semantics of these operations; we only need the machine to
distinguish them from normal execution. We model the machine as
a transition system in which some transitions are marked with
security operations. We then define a security semantics for stack safety
in terms of those steps.

The security semantics divides a program's execution into principals,
namely function activations, and identifies which principal is active
in a given state. Security operations can generate new principals, transition
which principal is active, and allocate stack objects belonging to the active
principal. From this information, we derive the specifications that make up
stack safety. Specifications may take the form of predicates on
future states: statements of the form, ``When control returns to me...'',
or of relations on traces of future execution (hyper-properties.)

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words (\(\WORDS\)) are 64-bit integers ranged over by \(\word\) and, when used as addresses,
\(\addr\). The processor has 32 registers, ranged over in general by \(\reg \in \REGS\),
plus a program counter (\(\PCname\)). The following table divides
the registers into metavariables, some belonging to caller- or callee- sets:

\vspace{\abovedisplayskip}
\begin{tabular}{| l | l | l |}
  \hline
  Set & Names & Purpose \\
  \hline
  CLR\_SV & \(rt_0 - rt_6\) & Caller-saved temps \\
  & \(ra_0 - ra_1\) & Caller-saved args / return vals \\
  & \(ra_2 - ra_7\) & Caller-saved args \\
  \hline
  CLE\_SV & \(rs_0 - rs_7\) & Callee-saved \\
  & \(rra\) & Return Address \\
  & \(rsp\) & Stack Pointer \\  
  \hline
  PUBLIC & \(pc\) & Program Counter \\
  \hline
\end{tabular}
\vspace{\abovedisplayskip}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to values.
The machine step function
\(\mach \xrightarrow{\psi} \mach' \in \MACHS \rightarrow \MACHS \times \Psi\)
may be labeled by a security operation, defined below.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by returning to the
same state. Since failstops are not distinguished from normal execution,
our properties will necessarily be {\it termination insensitive}.

@@

Security operations are split into two categories: {\it privilege operations}
and {\it control operations}. Privilege operations are those that adjust the
privileges of future callees by allocating private or public data and by saving
catch points for possible exceptions. Allocations specify the offset from
the stack pointer and size of the object. Exceptions are tied to the address
of the buffer that will contain saved state.
%
\begin{align*}
  \phi \in \Phi ::= & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{alloc\_pub} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}

Control operations are those that exchange
control between principals: calls (including tailcalls), returns, and thrown
exceptions. Calls and tailcalls identify the target of the call, and the argument registers.
[Still a todo: stack arguments.] Return needs no further information. Throwing
an exception specifies the address of the saved buffer.
%
\begin{align*}
  \chi \in \mathcal{X} ::= & \mathbf{call} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{tailcall} ~ \reg_{target} ~ \overline{\reg_{args}} & \\
  | & \mathbf{return} & \\
\end{align*}
%
A security operation may consist of any number of privilege operations and at most
one control operation.
%
\[\psi \in \Psi ::= \mathit{list ~ \Phi} | (\mathcal{X} \times \mathit{list ~ \Phi})\]

Security principals are function activations: at any given time, the current
activation, and some number of pending ones. Each activation will have a {\it view}
of the system that maps each element to one of a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \unsealed | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

These indicate, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
%in an allocated object that is accessible to the current principal (\(\mathit{object}\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).
The {\it initial view} \(V_0\) maps all stack locations and non-public registers
to \(\unsealed\) and all other locations and public registers to \(\public\).

@@

Every security operation manipulates the call stack: privilege operations act on
the active principal, and control operations push or pop the active principal.
We define the security effects of each operation as a function \(Pop\) or \(Cop\),
from an operation, a machine state, and a call stack to a new call stack.
As we introduce more complex operations, we will separately define
the curried functions \(Pop ~ \phi\) or \(Cop ~ \chi\) for each operation.
They are combined in \(Op\), which acts on full operations as follows.
\[Pop \in \Phi \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
\[Cop \in X \rightarrow \CONTEXTS \rightarrow \MACHS \rightarrow \CONTEXTS\]
\[Op ~ \psi =
\begin{cases}
  \lambda \context ~ \mach . \context & \textnormal{when } \psi = [~] \\
  Pop ~ \phi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = \phi::\overline{\phi} \\
  Cop ~ \chi \circ Op ~ \overline{\phi} & \textnormal{when } \psi = (\chi,\overline{\phi}) \\
\end{cases}\]

A control operation may additionally produce an observable event, drawn
from a set \(\mathit{EVENTS}\) and ranged over by \(e\).
A function \(Ev\) maps operations and machine states to events. The
silent event \(\tau\) is relevant to our definitions here, and the
remainder will be described in Section \ref{sec:events}. 
%
\[Ev \in X \rightarrow \MACHS \rightarrow \mathit{EVENTS}\]
%
From a machine state and a context stack, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with an event. If the underlying machine transition is labeled by
some operation \(\psi\), the combined step applies \(Op\) to the contents
of \(\psi\) to determine the new call stack; otherwise the call stack is
unchanged. If there is a control operation,
the transition is labeled with its event, otherwise it is silent.
Rules apply in order.

\judgmentthree[Control]
              {\(\mach \xrightarrow{\psi} \mach'\)}
              {\(\psi = (\chi, \overline{\phi})\)}
              {\(e = Ev ~ \chi ~ \mach\)}
              {\((\mach,\context) \stepstounder{e} (\mach',Op ~ \psi ~ \context ~ \mach)\)}

\judgment[NonControl]
         {\(\mach \xrightarrow{\psi} \mach'\)}
         {\((\mach,\context) \stepstounder{\tau} (\mach',Op ~ \psi ~ \context ~ \mach)\)}

@@

In this section we consider a simple setting with calls, returns, and private
allocations. Assume that \(\MACHS\) describe a RISC-V machine, where a transition
\(\mach \xrightarrow{\Psi?} \mach'\) is labeled as follows.
If \(\mach[\PCname]\) points to {\tt addi sp,sp,N}, \(\Psi? = \mathbf{alloc} ~ (0,N)\);
since we are not yet allocating public objects,
we don't need to distinguish objects from one another if they are allocated
simultaneously. Similarly, if \(\mach[\PCname]\) points to {\tt jal off ra}
and \(\mach'[\PCname] = a_f\) where \(a_f\) is the entry of a function
\(f\) that uses argument registers \(\overline{r}\), 
\(\Psi? = \mathbf{call} ~ f ~ \overline{r}\). And a step via the
instruction {\tt jalr ra} is labeled \(\mathbf{return}\).
This simplistic system serves to demonstrate our core properties.

Figure \ref{fig:example} gives a running example
of a function {\tt main} that takes a secret in its arguments and allocates
an array of potentially sensitive data. It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data.

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    {\tt
      void main(int argc, char*argv) \{

@@

      ~ ~ int sensitive[4] = \{0\};

      ~ ~ f();

      ~ ~ if (sensitive == 42) \{

@@

@@

      \}
    }
  \end{subfigure}
  \begin{subfigure}[t]{.5\textwidth}
    {\tt
      0: addi sp,sp,-20

      4: sw a1,16(sp)

      8: sd zero,8(sp)

      12: sd zero,0(sp)

      16: jal 84 ra

      20: lw a4,12(sp)

@@

      28: bne a4,a5,L

      32: ld a0,16(sp)

      36: jal 164 ra

      L(40): nop

      44: addi sp,sp,20

      48: return

    }
  \end{subfigure}

  \caption{Example: Main}
  \label{fig:example}
\end{figure*}

At the start of execution (PC 0) the initial call stack is \((V_0, [~])\).
Instruction 0 allocates 20 bytes starting at the stack pointer. This will have the
effect of marking those twenty bytes \(\sealed\), assuming they were previously
\(\unsealed\):

\judgmentbr{\(b = \mach[\SP] - \mathit{off}\)}
           {\(V' = V \llbracket \addr \mapsto \sealed |
             b \leq a < b+\mathit{sz} \land V ~ \addr = \unsealed \rrbracket\)}
           {\(Pop ~ (\mathbf{alloc} ~ \mathit{off, sz}) ~ (V,\sigma) ~ \mach \triangleq
             (V',\sigma)\)}

This memory being \(\sealed\) has no effect on the current activation --
the security guarantees apply at the start of the call. Our next security operation
occurs when we step from the call at instruction 16.

On a call, the formerly active principal's record is pushed onto the inactive list,
and replaced by a new record for the callee. Its return target is the instruction
immediately following that of the program counter before the call.
The callee's view is updated from the caller's such that the argument
registers are \(\public\) and any non-argument, caller-saved registers
are \(\unsealed\). All callee-save registers are always \(\sealed\), and
the program counter is always public, but this is already the case.

\[\begin{split}
\mathit{reg\_view} ~ V ~ \overline{r_{args}} \triangleq
V & \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR\_SV} \rrbracket \\
& \llbracket \reg \mapsto \public | \reg \in \overline{r_{args}} \rrbracket
\end{split}\]

\judgmenttwo{\(\chi = \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}}\)}
            {\(V' = \mathit{reg\_view} ~ V ~ \overline{\reg_{args}}\)}
            {\(Cop ~ \chi ~ (V,\sigma) ~ \mach \triangleq
              (V',(V, \mach[\PCname]+4, \mach[\SP])::\sigma)\)}

After the call step, we have a new activation, and {\tt main} is now pending.
{\tt main}'s pending record tracks its return address.
Everything that {\tt main} allocated, both {\tt secret} and {\tt sensitive},
{\tt f} views as \(\sealed\), along with its callee-saved registers.
Informally, the meaning of \(\sealed\) is: ``when control returns to
{\tt main}, this elements value should be the same as it is now.''
This is termed {\it return-time integrity}, and is defined below.
The remainder of the stack, and the caller-save registers, are \(\unsealed\).
The informal meaning of \(\unsealed\) is: ``{\tt f}'s behavior should not
depend on this element's value, but {\tt f} is free to overwrite it.''
This becomes relevant in out {\it confidentiality} properties.
Finally, \(\public\) elements are those that {\tt f} is allowed to
depend on, and may also overwrite.

On return, we restore the topmost inactive principal, if there is one.
Else, we default to retaining the current principal, as we do not model
the main function returning.

\judgmenttwo[Ret]
            {\(\chi = \mathbf{return}\)}
            {\(\sigma = (V,\addr_{ret},\addr_{sp})::\sigma'\)}
            {\(Cop ~ \chi ~ (\_, \sigma) ~ \mach \triangleq
              (V, \sigma)\)}

\judgment[RetDef]
         {\(\chi = \mathbf{return}\)}
         {\(Cop ~ \chi ~ (V, [~]) ~ \mach \triangleq
           (V, [~])\)}

We are ready to define our first stack safety property for this system:
{\it well-bracketed control flow}.

\section{Events and Traces}

Some macro-instructions may be interpreted as {\it events} that can be observed
from outside the system. For now we define events \(\obs \in \OBSS\) as calls, labeled
with the target address and the values of the arguments, or else as the silent
event \(\tau\):
\[\begin{split}
\obs \in \OBSS ::= & ~ \mathbf{callE} ~ a_{target} ~ \overline{w_{args}} \\
| & ~ \tau \\
\end{split}\]

We convert control operations into events via the \(Ev\)
function, observing the values of relevant registers, defined in Figure \ref{fig:observe}.

A {\em trace} is a nonempty, finite or infinite sequence
of events, ranged over by \(\obsT\).
We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
for list-cons.

\begin{figure*}
\[\mathit{Ev} ~ \chi ~ m =
\begin{cases}
  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \chi = \mathbf{call} ~ r_{target} ~ \overline{r_{args}} \\
  \mathbf{callE} ~ (m[\PCname]+m[\reg_{target}]) ~ m \llbracket r | r \in \overline{r_{args}} \rrbracket
  & \textnormal{when } ~ \chi = \mathbf{tailcall} ~ r_{target} ~ \overline{r_{args}} \\
  \tau & \textnormal{otherwise} \\
\end{cases}\]
\caption{Converting overlay steps to events}
\label{fig:observe}
\end{figure*}

@@

\judgmenttwo{\((\mach,\context) \stepstounder{e} (\mach',\context')\)}
            {\((\mach',\context') \hookrightarrow \obsT\)}
            {\((\mach,\context) \hookrightarrow \notfinished{e}{\obsT}\)}

@@

\section{Facts Abouts Calls and Returns}

>>>>

\section{The Setting}

We state our properties in terms of a generic machine subject to a few constraints,
but our examples and tests are focused on a RISC-V-like machine enhanced with PIPE,
a tag-based reference monitor. Importantly, such a machine has no explicit ``call''
instruction, only {\tt jal} instructions which may be used for that purpose or potentially
other purposes. Similarly, there is no ``return'', only {\tt jalr}, and so on.
We therefore define our properties under the assumption that we can distinguish the
steps that represent these abstract operations from those that do not -- an assumption
that is valid in our testing environment (Section [TBD]) and that is already necessary
to correctly apply tag-based enforcement in practice (cite something about this.)

The set of special operations that are recognized in this way are termed
{\it security operations}. Our properties are phrased as a {\it security semantics},
which extends the underlying machine with additional information about its
security principals and which registers and memory they require to be secure, which evolves
dynamically through the execution of security operations.
Specifications may take the form of
predicates on future states: statements of the form,
``When control returns to me...'', or of relations on traces of future execution
(hyper-properties).

This approach is in contrast to approaches that give a safe-by-construction semantics
for various security operations and then prove secure compilation to the underlying
enforcement mechanism, in that our machine is explicitly not safe unless the enforcement
is correctly instantiated; the security semantics gives us the means to distinguish an
enforcement scheme that implements stack safety from one that does not.

We will first give a detailed example of a security semantics for a simple setting
in which our security operations are restricted to making calls and returns, and allocating
private memory within the current function activation. We then extend this model
to the full system that we test, which features:
\begin{itemize}
\item Function calls and returns, with caller- and callee-saved registers
\item Allocation of strictly private and strictly public regions on the stack
\item Arguments spilled onto the stack and/or passed by reference
\item Exceptions
\item Tail-call Elimination
\end{itemize}

Finally, we give a property definition for stack safety in which (some) stack-allocated
objects are governed by a provenance-based capability system, accessible only to those
functions which have recieved a valid pointer to them.

In section (TBD) we discuss the extension of this model into a simple concurrency setup.

@@

We must trust that our method of distinguishing security operations is reasonable; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If it recognizes the operations based on the dynamic behavior of
the system, similarly, those definitions must be realistic.

Otherwise, we do not assume that the code is reasonable or corresponds to any particular
calling convention or source construct. In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism may place additional constraints, of course, particularly on the behavior of
call and return sequences.

In general, it is impossible to distinguish buggy source code from an attacker; in
our examples we will identify one function or another as an attacker, but we do not
require any static division between trusted and untrusted code, and aim to protect
even buggy code.

@@

Our concurrency model is fairly
simplistic, assuming a fixed number of threads each with its own dedicated processor.
We model memory safe stack objects, but not a heap. Regions outside of
stacks can be used however the compiler likes, including as a heap, but no protection is
built in and our properties assume that if a pointer to a stack object is stored there,
it is permanently compromised.

@@

The building blocks of the machine are {\em words} and {\em registers}.
Words are ranged over by \(\word\) and, when used as addresses, \(\addr\),
and are drawn from the set (\(\WORDS\)) (by default, the set of 64-bit integers).
Registers are ranged over in general by \(\reg \in \REGS\), other than the
program counter \(\PCname\) and stack pointer \(\SP\).
They are divided into sets of caller-saved and callee-saved registers;
Figure \ref{fig:RISCVregs} gives an example division for a RISC-V machine as will
appear in our examples and testing, along with the register names that will appear in
example code, and their default security class (explained below).

\begin{figure}
  \begin{tabular}{| l | l | l |}
    \hline
    Set / & Names & Purpose \\
    Class & & \\
    \hline
    \(\mathit{CLR}\) / & {\tt t0} -- {\tt t6} & Caller-saved temps \\
    \(\unsealed\) & {\tt a0} -- {\tt a1} & Caller-saved args / return vals \\
    & {\tt a2} -- {\tt a7} & Caller-saved args \\
    \hline
    \(\mathit{CLE}\) / & {\tt s0} -- {\tt s7} & Callee-saved \\
    \(\sealed\) & {\tt ra} & Return Address \\
    & {\tt sp} & Stack Pointer \\  
    \hline
    \(\mathit{PUBLIC}\) / & {\tt pc} & Program Counter \\
    \(\public\) & & \\
    \hline
  \end{tabular}
  \caption{RISC-V register set}
  \label{fig:RISCVregs}
\end{figure}

Collectively addresses and registers are {\em state elements} \(\component\)
in the set \(\COMPONENTS ::= \WORDS + \REGS\).
%
{\em Machine states} are drawn from a set \(\MACHS\) ranged over by \(\mach\)
that defines a mapping \(\mach[\component]\) from state elements to values.
{\em Events} are drawn from a set \(\OBSS\) and ranged over by \(\obs\).
The machine step function
\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \in \MACHS \rightarrow
\MACHS \times \mathit{list} ~ \Psi \times \OBSS\)
is labeled by a list of security operations (order matters), defined below,
and an event.
While the tag-based enforcement mechanism may cause the machine to
failstop, we model this as the machine silently diverging by remaining in the
same state. Since failstops are not distinguished from normal execution,
our properties will suffer the limitations typical of
{\it termination insensitive} systems \cite{}.

@@

Security operations consist of calls (including tailcalls), returns,
allocations and deallocations. Calls and tailcalls identify the address of the call's target,
and the argument registers. Allocations and deallocations are phrased in terms of a pair
of an offset and a size. Returns don't require any additional information.
%
\begin{align*}
  \psi \in \Psi ::= & \mathbf{call} ~ \addr_{target} ~ \overline{\reg_{args}} \\
  | & \mathbf{tailcall} ~ \addr_{target}  ~ \overline{\reg_{args}} \\
  | & \mathbf{return} & \\
  | & \mathbf{alloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
  | & \mathbf{dealloc} ~ \mathit{off ~ sz} & \mathit{off, sz} \in \mathbb{N} \\
\end{align*}
%
Security principals are function activations: at any given time, the current
activation, and some number of pending ones. Each activation will have a {\it view}
of the system that maps each element to one of a set of {\it security classes}:
\[sc \in SEC ::= \sealed | \unsealed | \object | \public\]
\[V \in \mathit{VIEW} ::= \COMPONENTS \rightarrow SEC\]

These indicate, from the perspective of a given principal, whether an element is
private to an inactive principal (\(\sealed\)),
in an allocated object that is accessible to the current principal (\(\object\)),
available to be allocated (\(\unsealed\)), or
always accessible (\(\public\)).
The {\it initial view} \(V_0\) maps all stack locations to \(\unsealed\),
all other locations to \(\public\), and registers based on which set they
belong to (Figure \ref{fig:RISCVregs}): \(\sealed\) for callee-saved,
\(\unsealed\) for caller-saved, and \(\public\) otherwise.

@@

Every security operation manipulates the call stack via a function
\(Op : \Psi \rightarrow \MACHS \rightarrow \CONTEXTS \rightarrow \CONTEXTS\). We will introduce the
curried application of \(Op\) to each operation as we introduce them in detail.

From a machine state and a context stack, we create a {\it combined state}
\((\mach,\context)\) and a transition \(\stepstounder{}\) on combined states,
labeled with the same operations. Note that \(Op\) takes as its state argument
\(\mach\), the state before the step.

\judgmenttwo{\(\mach \xrightarrow{\overline{\psi}} \mach', \obs \)}
            {\(\mathit{foldl} ~ (Op ~ \psi ~ \mach) ~ \context = \context'\)}
            {\((\mach,\context) \stepstounder{\psi :: \overline{\psi}} (\mach', \context'', \obs)\)}

@@

In this section we consider a simple setting with calls, returns, and private allocations.
Figure \ref{fig:main} gives a C function {\tt main}, and its compiled RISC-V code,
that takes a secret in its arguments and allocates
an array of potentially sensitive data. It then calls another function {\tt f},
and afterward may decide to {\tt publish} its secret based on the contents of
its sensitive data.

\begin{figure}
  \begin{subfigure}[b]{\columnwidth}
    {\tt
      void main(int argc, int* argv) \{

@@

      ~ ~ int sensitive[4] = \{0,0,0,0\};

      ~ ~ int res = f();

      ~ ~ if (sensitive[0] == 42) \{

@@

      ~ ~ \} else \{

      ~ ~ ~ ~ publish(res);

@@

      \}}
  \end{subfigure}
  \begin{subfigure}[b]{.45\columnwidth}
    \vspace{2\abovedisplayskip}

    {\tt
      0: addi sp,sp,-24

      4: sw a1,24(sp)

      8: sd zero,16(sp)

      12: sd zero,8(sp)

      16: jal 84,ra

      20: lw a4,4(sp)

@@

      28: bne a4,a5,L1

      32: lw a0,16(sp)

      36: jal 164,ra

      40: j L2:

      L1:

      44: jal 164,ra

      L2:

      48: addi sp,sp,24
    }
  \end{subfigure}
  \begin{subfigure}[b]{.5\columnwidth}
    \includegraphics[width=\columnwidth]{stacklayout.png}
  \end{subfigure}

\caption{Example: Main}
\label{fig:main}
\end{figure}

Suppose that {\tt f} is not a (valid) C function at all, but an attacker seeking
to leak {\tt secret}! It might do so in a number of ways, shown as snippets of
assembly code in Figure \ref{fig:f}. It might take an offset from the stack
pointer, access {\tt secret}, and directly call {\tt publish} itself. But more
subtly, even if somehow prevented from outputting {\tt secret} directly, it might
instead return that value so that {\tt main} makes the call to {\tt publish}.

Beyond simply reading {\tt secret}, the attacker might overwrite {\tt sensitive[0]}
with 42, guaranteeing that {\tt main} publishes its own secret unintentionally!
Attacks of this kind do not violate {\tt main}'s confidentiality, but its
{\it integrity}.
And last, but certainly not least, the attacker might attempt to return to the
wrong instruction, thereby bypassing the check and publishing {\tt secret} regardless,
violating the program's {\it well-bracketed control flow} (WBCF.)

\begin{figure}
  \begin{subfigure}[b]{.5\columnwidth}
    {\tt
      100: addi sp,sp,-4

      104: sw ra,4(sp)

      108: lw a0,20(sp)

      112: jal 88,ra

      116: lw ra,4(sp)

      120: jalr ra
    }
    \caption{Leaking {\tt secret} directly}
  \end{subfigure}  
  \begin{subfigure}[b]{.45\columnwidth}
    {\tt
      100: lw a4,16(sp)

      104: sw a4,12(sp)

      108: jalr ra
    }
    \caption{Leaking {\tt secret} indirectly}
  \end{subfigure}

  \vspace{\belowdisplayskip}

  \begin{subfigure}[b]{.5\columnwidth}
    {\tt
      100: li a5,42

      104: sw a5,12(sp)

      108: jalr ra
    }
    \subcaption{Attacking {\tt sensitive}}
  \end{subfigure}
  \begin{subfigure}[b]{.45\columnwidth}
    {\tt
      100: addi ra,ra,20

      104: jalr ra
    }
    \subcaption{Attacking control flow}
  \end{subfigure}

  \caption{Assembly code for {\tt f} as an attacker}
  \label{fig:f}
\end{figure}

At the start of execution the program counter is 0 and the stack pointer -4, with code
proceeding upward and the stack growing down. The functions {\tt f} and {\tt publish}
are at addresses 100 and 200, respectively.

Figure \ref{fig:exec1} shows how the program counter, stack pointer, and call stack
develop over several steps, with labels on the steps that allocate space and then make
a call. \(\downarrow \overline{\psi}\) in this diagram is a sideways
\(\xrightarrow{\overline{\psi}}\). Events are omitted.

The first step allocates a word for {\tt secret} and four for {\tt sensitive}. To
maintain a two-word alignment it leaves a gap. This will have the
effect of marking those bytes \(\object\), assuming they were previously
\(\unsealed\).

\begin{figure*}
  \begin{subfigure}[t]{.5\textwidth}
    \vskip 0pt
    \begin{tabular}{| l | l | l | l |}
      \hline
      Step & \(\PCname\) & \(\SP\) & \(\context\) \\
      \hline
      0 & 0 & 0 & \(V_0, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{alloc} ~ (-4,4);\mathbf{alloc} ~ (-24,16)]\)} \\
      \hline
      1 & 4 & -24 & \(V_1 = V_0 \llbracket -24 \ldots -9, \mapsto \object \rrbracket\) \\
          & & & \(\llbracket -4 \ldots -1, \mapsto \object \rrbracket, \emplist\)\\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      4 & 16 & -24 & \(V_1, \emplist\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{call} ~ 100 ~ \emplist ~ \emplist]\)} \\
      \hline
      5 & 100 & -24 & \(V_2 = V_1 \llbracket \reg \mapsto \unsealed | \reg \in \mathit{CLR} \rrbracket,\) \\
      & & & \([(V_1, 20, -24)]\) \\
      \hline
      \multicolumn{4}{l}{\(\vdots\)} \\
      \hline
      \(n\) & ? & ? & \(V_2,[(V_1, 20, -24)]\) \\
      \hline
      \multicolumn{4}{l}{\(\downarrow [\mathbf{return}]\)} \\
      \hline
      \(n+1\) & 20 & -24 & \(V_1, \emplist\) \\
      \hline
    \end{tabular}
  \end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
    \vskip 0px
    \includegraphics[width=\columnwidth]{stack1.png}
    \includegraphics[width=\columnwidth]{stack2.png}
    \includegraphics[width=\columnwidth]{stack3.png}
    \includegraphics[width=\columnwidth]{stack4.png}
  \end{subfigure}

\caption{Execution through return from {\tt f}}
\label{fig:exec1}
\end{figure*}

On a call, the formerly active principal's record is pushed onto the inactive list.
Its return target is the instruction immediately following that of the program
counter before the call, and the stack pointer target is the current stack pointer.
The callee's view is updated from the caller's such that the argument
registers are \(\public\) and any non-argument, caller-saved registers
are \(\unsealed\). All callee-save registers are always \(\sealed\) and
the program counter \(\public\), as they were initialized.

We suppose that, on step \(n+1\), {\tt f} returns; if it does not, then
{\tt main}'s secrets are safe as long as {\tt f} does not leak them directly.
On that return, we restore the topmost inactive principal, {\tt main}'s principal.

\paragraph*{Well-bracketed Control Flow}

But what if {\tt f}, does not return with \(\PCname = 20\) and \(\SP = -24\),
as would be the case in the fourth example attack? Then it has violated WBCF.
This property can be phrased in terms of a single step: for every transition
labeled as a return, the program counter and stack pointer match those recorded
in the top pending activation.

\paragraph*{Stack Integrity}

Integrity is more complicated, and requires us to compare the state at the start of
the call to {\tt f} to that on return. Intuitively, the integrity of {\tt main}
is preserved if, when control returns to it, it can rely on any \(\sealed\) elements
to be identical to when it made the call.

In the case of the call from {\tt main} to {\tt f}, the \(\sealed\) elements are the
addresses -24 through -9 and -4 through -1 (allocated as private data) and the callee-saved registers.
Note that the callee-saved registers may well have changed during the call -- but if
so, the callee is obligated to restore them to their prior value.

One further caveat: it is possible that {\tt f} might overwrite {\tt main}'s data
in a way that is harmless -- or that is rendered harmless by the enforcement mechanism!
This is the case in a lazy micro-policy: the monitor permits the caller to write into the
callee's frame, but taints that write so that the callee cannot read it later without
triggering a failstop. In such a scenario we will consider a value {\it harmless}
if it could be replaced with any other value without changing the observable behavior
of the machine. That is, for a given state \(\mach\) and set of elements \(\components\),
we define a {\it variant} state that agrees with \(\mach\) on every element except those
in \(\components\). Those elements are said to be harmless in \(\mach\) if the variant
state produces an equivalent trace (see Section \ref{sec:events}.)

So, to define integrity, we need to know when a caller has been returned to;
this is the first state after the call in which the depth of the call stack
matches the state before the call. Integrity means that, at that return state,
any element that is \(\sealed\) under the callee's view is either restored
to its original value or is harmless.

\paragraph*{Caller Confidentiality}

We treat confidentiality as a form of non-interference as well: the confidentiality of a caller
means that its callee's behavior is dependent only on publicly visible data,
not the caller's state. As we saw in our examples, the callee's behavior is defined
both in terms of events that the callee produces during the call, but also in terms of
the changes it makes to the state.

Consider the state \((\mach, (V_2,\sigma))\) at step 5. If we take a variant state over
the set of element(s) that are \(\sealed\) or \(\unsealed\) in \(V_2\),
{\it internal confidentiality} means
that, if we execute both states until they return, they will produce equivalent sequences
of events.

But, it is also possible that the callee could extract data from the caller that it does
not use to produce a visible event, but instead stashes for later or uses to influence
its return value -- which in turn can lead to future changes in behavior that do not
even appear to be the fault of the callee!
{\it Return-time confidentiality} means that any state element that changes in one variant
must either change in both, or be harmless.

Together, these definitions encompass the confidentiality of the caller from the callee.

\paragraph*{Callee Confidentiality}

Although we presented our initial example from the perspective of the caller, a callee
may also have data that must be kept secret from its caller. Consider a callee that makes
a privileged system call to obtain a secret key, and uses that key to perform a specific
task. An untrustworthy or erroneous caller might attempt to read the key out of the callee's
memory after return.

We model this as nearly dual to caller integrity: the callee's secrets are those
values that it has stored in \(\unsealed\) elements. So, callee confidentiality means that
at the state following the callee's return, any element that was \(\unsealed\) under the
callee's view either retains its original value or is harmless.

\paragraph*{The Hierarchy of Stack Safety}

The forms of stack safety described above are largely independent of one another, and
can potentially be enforced separately in various combinations. This is important, because
many enforcement mechanisms already in existence only enforce some of the stack safety
properties. Stricter enforcement may have greater performance impacts, leading to decisions
about trade-offs between different points in a hierarchy of properties.

At the most basic, if we protect nothing else, it is vital to protect the control flow of the
system by enforcing WBCF. None of the data-protection properties are meaningful if a
callee can simply ``return'' to an arbitrary instruction and proceed with the privileges
of its caller! And from a practical standpoint, control-flow attacks such as return-oriented
programming are among the most severe attacks involving the stack and the core motivation
for much existing work.

\paragraph*{Moving Forward}

In the next section we will define traces and their (hyper-)properties.
Then in Section \ref{sec:facts} we will define a mechanism to make assertions
about what will happen on a future return, and Section \ref{sec:props} will use these
to give the formal definitions of integrity and confidentiality.

\section{Events and Traces}
\label{sec:events}

We abstract over the events that can be observed in the system, defining them
only as a set \(\OBSS\) that contains at least the element \(\tau\), the silent
event. Other events might represent certain function calls (i.e., system calls)
or writes to special addresses representing mmapped regions.
A {\em trace} is a nonempty, finite or infinite sequence
of events, ranged over by \(\obsT\).
We use ``\(\notfinished{}{}\)'' to represent ``cons'' for traces, reserving ``::''
for list-cons.

@@

\judgmenttwo{\((\mach,\context) \stepstounder{} (\mach',\context',\obs)\)}
            {\((\mach',\context') \hookrightarrow \obsT\)}
            {\((\mach,\context) \hookrightarrow \notfinished{\obs}{\obsT}\)}

@@

\paragraph*{Variants and Harmless Values}

Two states are variants with respect to a set of elements, \(\components\),
if they agree on the value of every element not in \(\components\). Our
notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of vestigial elements.
